The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-06 analysis/paper/paper.qmd
Shorten `emu_theses` to `emu` in all object names

## Read only metadata for theses with PDF available
emu_theses <- 
  read_csv(here(data_path, "theses-metadata.csv")) |>
  mutate(text = "") |>
  filter(!is.na(pdf_via))

pdf_paths <- here(data_path, emu_theses$file_name)
```

@@

if (!file.exists(data_path)) {
  # On the first run, this function might take extra time as a Python environment needs to be set up
  emu_theses$text <- convert_pdf_text(pdf_paths)

  # Add latitude and longitude information to theses data frame
  emu_theses <- geocode_thesis_locations(emu_theses)

  # Write thesis dataframe to file
  write_csv(emu_theses, data_path)
} else {
  emu_theses$text <- read_csv(data_path)$text
  emu_theses$longitude <- read_csv(data_path)$longitude
  emu_theses$latitude <- read_csv(data_path)$latitude
}

@@

emu_theses$text_clean <- emu_theses$text |>
  clean_basic()

@@

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program. To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.

@@

A total number of `r nrow(emu_theses)` out of `r nrow(all_theses)` were analysed.

@@

# Show thesis locations on world map
visualize_thesis_locations(emu_theses)
```

@@

# Add cleaned text column with removed stop words
emu_theses$text_clean <- emu_theses$text_clean |>
  str_remove_all(find_meta_stopwords(emu_theses)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

# Unnest words and remove words of one and two letters
emu_theses_words <- emu_theses |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
  mutate(word = textstem::lemmatize_words(word))  # lemmatise words

@@

#short_words(emu_theses_words$word, l_word = 3, l_list = 100)

emu_theses_words <- emu_theses_words |>  
  filter(nchar(word) >= 3)  # keep only words that are at least three letters long

@@

# Prepare document-term matrix as input for LDA model
emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")

@@

# The top n most common words in each topic
emu_theses_top_terms <- get_top_words_per_topic(emu_topics, 20)

# TODO create visualisation function
emu_theses_top_terms |> 
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

# TODO can this be generalised to more than two topics?
beta_spread <- emu_theses_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

@@

emu_documents |> 
  inner_join(emu_theses |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_documents |> 
  inner_join(emu_theses |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```

@@

```{r top-words-document}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_theses_words, top_n = 3)
```

```{r top-words-corpus}
# Which are the top n most frequently used words in all theses?
# TODO add `topic` to `fill` aesthetic
n = 50
get_top_words_per_corpus(data = emu_theses_words, top_n = n) |>
  ggplot(aes(reorder(word, n), n)) +
  geom_col(fill = "orange") +
  geom_text(aes(label = n), size = 2, hjust = 1.1) +
  coord_flip() +
  xlab("Word") +
  ylab("Frequency") +
  labs(title = paste0("Top ", n, " most used words in the corpus of theses")) +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

@@

## Bigrams
emu_bigrams <- emu_theses |> 
  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)

@@

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text")

trigram_counts <- emu_theses_trigrams |> 
  count(w_1, w_2, w_3, sort = TRUE)

@@

## Trigram tf-idf
trigram_tf_idf <- emu_theses_trigrams |> 
  count(title, ngram) |> 
  bind_tf_idf(ngram, title, n) |> 
  arrange(desc(tf_idf))

@@

## Calculate tf-idf statistic
emu_theses_tf_idf <- emu_theses_words |> 
  get_tf_idf(title_col = "title", word_col = "word")

## Get top n words most specific to each thesis 
emu_theses_tf_idf |> 
  group_by(title) |> 
  top_n(5, tf_idf) |> 
  arrange(desc(tf_idf), .by_group = TRUE)

@@

```{r}
# Preview term frequency distributions in a subset of theses
ggplot(emu_theses_words_count |> 
         filter(title == unique(title)[1:6]),
       aes(n/total, fill = title)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ title, ncol = 3, scales = "free_y")
```

```{r}
# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_theses_words_count |> 
  group_by(title) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

@@

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_theses_words_count |> 
  filter(title == unique(title)[1:6]) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

@@

# To what extent do words co-occur with other words?
tidy_pmi <- get_pmi(emu_theses, title_col = "title", text_col = "text_clean")

@@

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_emu_theses |> 
  count(title, word) |> 
  cast_sparse(title, word, n)

>>>>

## Read only metadata for theses with PDF available
emu <- 
  read_csv(here(data_path, "theses-metadata.csv")) |>
  mutate(text = "") |>
  filter(!is.na(pdf_via))

pdf_paths <- here(data_path, emu$file_name)
```

@@

if (!file.exists(data_path)) {
  # On the first run, this function might take extra time as a Python environment needs to be set up
  emu$text <- convert_pdf_text(pdf_paths)

  # Add latitude and longitude information to theses data frame
  emu <- geocode_thesis_locations(emu)

  # Write thesis dataframe to file
  write_csv(emu, data_path)
} else {
  emu$text <- read_csv(data_path)$text
  emu$longitude <- read_csv(data_path)$longitude
  emu$latitude <- read_csv(data_path)$latitude
}

@@

emu$text_clean <- emu$text |>
  clean_basic()

@@

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program. To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.

@@

A total number of `r nrow(emu)` out of `r nrow(all_theses)` were analysed.

@@

# Show thesis locations on world map
visualize_thesis_locations(emu)
```

@@

# Add cleaned text column with removed stop words
emu$text_clean <- emu$text_clean |>
  str_remove_all(find_meta_stopwords(emu)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

# Unnest words and remove words of one and two letters
emu_words <- emu |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
  mutate(word = textstem::lemmatize_words(word))  # lemmatise words

@@

#short_words(emu_words$word, l_word = 3, l_list = 100)

emu_words <- emu_words |>  
  filter(nchar(word) >= 3)  # keep only words that are at least three letters long

@@

# Prepare document-term matrix as input for LDA model
emu_dtm <- convert_to_dtm(emu_words, "title", "word")

@@

# The top n most common words in each topic
emu_top_terms <- get_top_words_per_topic(emu_topics, 20)

# TODO create visualisation function
emu_top_terms |> 
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

# TODO can this be generalised to more than two topics?
beta_spread <- emu_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

@@

emu_documents |> 
  inner_join(emu |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_documents |> 
  inner_join(emu |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```

@@

```{r top-words-document}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_words, top_n = 3)
```

```{r top-words-corpus}
# Which are the top n most frequently used words in all theses?
# TODO add `topic` to `fill` aesthetic
n = 50
get_top_words_per_corpus(data = emu_words, top_n = n) |>
  ggplot(aes(reorder(word, n), n)) +
  geom_col(fill = "orange") +
  geom_text(aes(label = n), size = 2, hjust = 1.1) +
  coord_flip() +
  xlab("Word") +
  ylab("Frequency") +
  labs(title = paste0("Top ", n, " most used words in the corpus of theses")) +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

@@

## Bigrams
emu_bigrams <- emu |> 
  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)

@@

## Trigrams
emu_trigrams <- emu |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text")

trigram_counts <- emu_trigrams |> 
  count(w_1, w_2, w_3, sort = TRUE)

@@

## Trigram tf-idf
trigram_tf_idf <- emu_trigrams |> 
  count(title, ngram) |> 
  bind_tf_idf(ngram, title, n) |> 
  arrange(desc(tf_idf))

@@

## Calculate tf-idf statistic
emu_tf_idf <- emu_words |> 
  get_tf_idf(title_col = "title", word_col = "word")

## Get top n words most specific to each thesis 
emu_tf_idf |> 
  group_by(title) |> 
  top_n(5, tf_idf) |> 
  arrange(desc(tf_idf), .by_group = TRUE)

@@

```{r}
# Preview term frequency distributions in a subset of theses
ggplot(emu_words_count |> 
         filter(title == unique(title)[1:6]),
       aes(n/total, fill = title)) +
  geom_histogram(show.legend = FALSE, binwidth = 0.0001) +
  xlim(NA, 0.0009) +
  facet_wrap(~ title, ncol = 3, scales = "free_y")
```

```{r}
# Examine Zipf's law for the EMU theses
## Calculate rank and term frequency
freq_by_rank <- emu_words_count |> 
  group_by(title) |> 
  mutate(rank = row_number(),
         `term frequency` = n/total)

@@

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_words_count |> 
  filter(title == unique(title)[1:6]) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

@@

# To what extent do words co-occur with other words?
tidy_pmi <- get_pmi(emu, title_col = "title", text_col = "text_clean")

@@

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_emu |> 
  count(title, word) |> 
  cast_sparse(title, word, n)
